---
output: github_document
---

[![Travis-CI Build Status](https://travis-ci.org/SymbolixAU/RcppViridis.svg?branch=master)](https://travis-ci.org/SymbolixAU/RcppViridis)
[![Coverage status](https://codecov.io/gh/SymbolixAU/RcppViridis/branch/master/graph/badge.svg)](https://codecov.io/github/SymbolixAU/RcppViridis?branch=master)
[![Github Stars](https://img.shields.io/github/stars/SymbolixAU/RcppViridis.svg?style=social&label=Github)](https://github.com/SymbolixAU/RcppViridis)

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#",
  fig.path = "man/figures/README-",
  out.width = "100%",
  out.height = 200
)

library(RcppViridis)
```

# RcppViridis

---

### What does it do?

It maps viridis colours to variables, and quickly!

### Why did you build it?

I'm aware there are other methods for mapping colours to values. And which do it quick too. But I can never remember them, and I find the interfaces a bit cumbersome. For example,  `scales::col_numeric(palette = viridisLite::viridis(5), domain = range(1:5))(1:5)`.

I wanted **one** function which will work on **one** argument. 

```{r}
colour_values(1:5)
colour_values(letters[1:5])
```

I also want it available at the `src` (C/C++) level for linking to other packages. 

---

### Why do you spell colour with a 'u'?

Because it's correct, and [R tells us to](http://developer.r-project.org/Rds.html)

> For consistency, aim to use British (rather than American) spelling

But don't worry, `color_values(1:5)` works as well

---

### How do I install it?

Install the development version from [GitHub](https://github.com/SymbolixAU/RcppViridis) with:

```r
# install.packages("devtools")
devtools::install_github("SymbolixAU/RcppViridis")
```

---

### How can I make use of it in my package?

**Rcpp**

All functions are written in `Rcpp`. I have exposed some of them in header files so you can `LinkTo` them in your package.

**R**

If you're not using `Rcpp`, just `Import` this package like you would any other.

## Do you have any examples?

Of course!


#### 256 numbers mapped to a colour

```{r}
df <- data.frame(a = 10, x = 1:256)
df$col <- colour_values(df$x, palette = "viridis")
barplot(height = df$a, col = df$col, border = NA, space = 0)
```

#### 5000 numbers on a non-linear scale

```{r}
df <- data.frame(a = 10, x = c((1:5000)**3))
df$col <- colour_values(df$x, palette = "viridis")
barplot(height = df$a, col = df$col, border = NA, space = 0)
```

#### 1000 random numbers

```{r}
df <- data.frame(a = 10, x = rnorm(n = 1000))
df$col <- colour_values(df$x, palette = "inferno")
barplot(height = df$a, col = df$col, border = NA, space = 0)
```

Eurgh!

```{r}
df <- df[with(df, order(x)), ]
barplot(height = df$a, col = df$col, border = NA, space = 0)
```

That's better!


---

## Do I have to use the in-built palettes?

No, you can use your own specified as a matrix of red, gree, blue columns in the range [0,1]

```{r}
n <- 100
m <- grDevices::colorRamp(c("red", "green"))( (1:n)/n )
df <- data.frame(a = 10, x = 1:n)
df$col <- colour_values(df$x, palette = m)
barplot(height = df$a, col = df$col, border = NA, space = 0)
```


## Do you support 'alpha' values

Yep. Either supply a single alpha value for all the colours

```{r}
## single alpha value for all colours
df <- data.frame(a = 10, x = 1:255)
df$col <- colour_values(df$x, alpha = 50)
barplot(height = df$a, col = df$col, border = NA, space = 0)
```

Or include the alpha value as a 4th column in the palette matrix

```{r}
n <- 100
m <- grDevices::colorRamp(c("red", "green"))( (1:n)/n )
## alpha values
m <- cbind(m, seq(0, 255, length.out = 100))
df <- data.frame(a = 10, x = 1:n)
df$col <- colour_values(df$x, palette = m)
barplot(height = df$a, col = df$col, border = NA, space = 0)
```

---

### What's the performance like?



**1 million numeric values**

```{r, out.height=400}
library(microbenchmark)
library(ggplot2)
library(scales)
library(viridisLite)

n <- 1e6
df <- data.frame(x = rnorm(n = n))

m <- microbenchmark(
  RcppViridis = { RcppViridis::colour_values(x = df$x) },
  scales = { col_numeric(palette = rgb(subset(viridis.map, opt=="D")[, 1:3]), domain = range(df$x))(df$x) },
  times = 25
)
m

autoplot(m)
```


**1 million characters (26 unique values)**

```{r, out.height=400}
library(microbenchmark)
library(ggplot2)
library(scales)
library(viridisLite)

n <- 1e6
x <- sample(x = letters, size = n, replace = TRUE)
df <- data.frame(x = x)

m <- microbenchmark(
  RcppViridis = { x <- RcppViridis::colour_values(x = df$x) },
  scales = { y <- col_factor(palette = rgb(subset(viridis.map, opt=="D")[, 1:3]), domain = unique(df$x))(df$x) },
  times = 25
)
m

autoplot(m)
```

